---
title: "HR Attrition"
output: html_notebook
---

# Introduction:

In this notebook I will be looking at IBM's employee attrition data obtained from Kaggle https://www.kaggle.com/pavansubhasht/ibm-hr-analytics-attrition-dataset

My goal in this notebook is to predict the likeklihood of employees leaving teh company and assessing the impact this will have on the company.
```{r}
library(tidyverse)
library(caret)
library(DataExplorer)
setwd("~/HRattrition")
df = read.csv("WA_Fn-UseC_-HR-Employee-Attrition 2.csv")
df$Over18 = NULL
df$EmployeeCount = NULL
df$StandardHours = NULL
```

# EDA
There are 35 variables and 1470 observations of these in this dataframe.  I will use the head() command to display the first few rows
```{r}
#taking a quick look at the data 
df %>% head()
```

The plot_histogram() function from DataExplorer is useful to plot the distributions of all numeric values in a dataframe.  One thing to note is that Age is the only variable that is nearing normality.

```{r}
df %>% plot_histogram()
```

There is no missing values in the whole dataset.  This is helpful because missing values are problematic when modeling.  

```{r}
df %>% plot_missing()
```

A quick correlation matrix of the numeric features higlights that WorklifeBalance, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion and YearsWithCurrentManager contain much of the same information and are highly correlated.

```{r}
df %>% select_if(is.numeric) %>%  plot_correlation(type = "c")
```

The varible of interest for prediction is _Attrition_ a 2 level factor that indicated whether turnover occured with the given observation.  I plot a simple barplot of the Attrition variable.


```{r}
df %>% ggplot(aes(Attrition)) +
  geom_bar()+
  ggtitle("Employee Turnover")
```
# Modeling 

The first model that I fit is a random forest model.  

```{r}
library(caret)
split = createDataPartition(df$Attrition, p = 0.7, list = F)
train = df[split,] %>% drop_na()
test = df[-split,] %>% drop_na()
```



```{r}
library(randomForest)
rf =randomForest::randomForest(Attrition~., data = train)
pred = predict(rf, test, type = "response")
#pred[pred > 0.5] = "Yes"
#pred[pred < 0.5] = "No"
mean(pred == test$Attrition)
confusionMatrix(pred, test$Attrition)
```
```{r}
rf
```

```{r}
pred2 = predict(rf, test, type = "prob")
data.frame(pred2) %>% select(No) %>% 
ggplot(aes(No)) + geom_histogram(bins=50) + ggtitle("Probability of Staying Based on Random Forest")

```


```{r}
varImpPlot(rf, n.var = 17)
```

```{r}
df$Attrition = as.factor(df$Attrition)
```


```{r}
library(pdp)
rf %>%
  partial(pred.var = "DailyRate") %>%
  plotPartial(rug = TRUE, train = train, main = "Monthly Income Partial Dependence Plot")
```

```{r}
library(pdp)
rf %>%
  partial(pred.var = "Age") %>%
  plotPartial(rug = TRUE, train = train, main = "Age Partial Dependence Plot")
```
```{r}
library(pdp)
rf %>%
  partial(pred.var = "OverTime") %>%
  plotPartial(rug = TRUE, train = train)
```






```{r}
ctrl <- trainControl(method = "repeatedcv", repeats = 3)
regFit <- train(
  Attrition ~ .,
  data = train %>% drop_na(),
  trControl = ctrl,
  method = "glm",
  #family = "binomial",
  ## Center and scale the predictors for the training
  ## set and all future samples.
  preProc = c("center", "scale"),
  tuneLength = 10
)
```

```{r}
pred = predict(regFit, test, type = "prob")
pred2 = predict(regFit, test)
confusionMatrix(pred2, test$Attrition)
```

```{r}
data.frame(pred) %>% select(No) %>% ggplot(aes(No))+ 
  geom_histogram(bins=50)
```


```{r}
library(MASS)
coef = add_rownames(data.frame(linearmod$coefficients), "coefficent")
coef %>% ggplot(aes(coefficent,linearmod.coefficients))+
  geom_bar(stat="identity")+ labs(y="Effect") +
  theme(axis.text.x = element_text(size = 8, angle = 90))
```

```{r}
as.factor(pred)
```



